---
title: "Changepoints"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# install.packages('pacman')
library(pacman)
pacman::p_load(
    tidyverse,
    scales,
    lubridate,
    reshape2, 
    patchwork
)

```

## Data and setup
Loading the data
```{r}
df <- read_csv("idmdl/changepoints/tweets_emo_date_W3_cp_4.csv") %>%
                mutate(change_point=change_point+1) #%>%
                # filter(resonance < 0.01 & novelty < 0.01)

```


Preparing datecolumn
```{r}
df <- df %>% mutate(date = as.Date(date))
```

See dates in change point windows
```{r}
head(df)
df %>% group_by(change_point) %>% summarise(min_date = min(date),
                                             max_date = max(date))
```


Define function
```{r}
# For signal including 2022
get_text_position <- function(change_point){
    y = 0.0055
    if (change_point == 1){
        x = as.Date("2019-07-01")
    }
    else if (change_point == 2){
        x = as.Date("2020-07-01")
    }
    else if (change_point == 3){
        x = as.Date("2020-12-28")
        y = 0.011
    }
    else if (change_point == 4){
        x = as.Date("2021-07-01")
    }
    else if (change_point == 5){
        x = as.Date("2021-12-28")
        y = 0.011
    }
    else if (change_point == 6){
        x = as.Date("2022-04-20")
    }
    return(c(x, y))
}
```


Prepare dfs for coloring background by change point and adding text
```{r}
# # Remove if they exist before loop
rm(color_df)
rm(text_df)

###########################################
# Background color and change point title #
###########################################

for (point in unique(df$change_point)){ # Loop over change points
  # Subset from df
  sub = df %>% subset(change_point == point)
  # Get text position x and y
  positions = get_text_position(point)
  
  # Df for background color
  temp_col <- tibble(start = min(sub$date)-1,
                 end = max(sub$date),
                 change_point = point)

  # Df for change point title text
  temp_text <- tibble(x_pos = positions[1],
                      y_pos = as.numeric(positions[2]),
                      text = paste("Change point period", point, sep = " "),
                      change_point=point)
  
  # Bind
  if (exists('color_df')){color_df <- rbind(color_df, temp_col)} else {color_df <- temp_col}
  if (exists('text_df')){text_df <- rbind(text_df, temp_text)} else {text_df <- temp_text}
}

# Make background go all the way 
color_df$start[1] = -Inf
color_df$end[6] = Inf

# Remove temporary dfs from environment
rm(temp_col, temp_text)


######################################
# Manually define events of interest #
######################################
covid_manual <- df %>% 
    subset(date == "2020-01-30" | # WHO: covid is a threat
           date == "2020-02-26" | # First DK citizen
           date == "2020-03-11" | # Lockdown
           date == "2020-03-17" | # Queen's speech
           date == "2020-04-20" | # Partial reopening
           date == "2020-07-31" | # face masks
           date == "2020-11-04" | # kill the mink
           date == "2020-12-27" | # First dk vaccination
           date == "2021-01-28" | # Lockdown prolonged
           date == "2021-04-14" |  # AZ withdrawn
           date == "2021-05-28" | # Coronapas-app published in App-store
           date == "2021-07-05" | # delta variant dominat
           date == "2021-09-01" | # no longer "samfundskritisk sygdom"
           date == "2021-11-11" | # again "samfundskritisk sygdom"
           date == "2022-02-01" | # no longer "samfundskritisk sygdom"
           date == "2022-04-06"  # quick testcenters closes
            ) %>% 
    mutate(change_point = as.factor(change_point),
           n_event = seq(1,16),
           date = ifelse(n_event==9, as.Date(date)+4,ifelse(n_event==10, as.Date(date)-10, as.Date(date))),
           y_pos= ifelse(resonance>=0, resonance+0.001, resonance-0.0008)) %>%
    mutate(date= as.Date(date, origin="1970-01-01"))

```

Defining colors

```{r}
color1 = "#2C76A8"
color2 = "#4cbbc2" # "#4CC5CB"
color3 = "#004079"
color4 = "#008C76FF"

fill_colors= c(color1, color2, color3, color4, color1, color2)
col_colors = c(color1, color2, color3, color4, color1, color2, color2, color3, color4)

```

# Raw resonance and novelty
```{r}
res_plot <- df %>% ggplot(aes(y=resonance, x=date)) +
                geom_line(color = "steelblue") +
                labs(x = "", y = "Resonance") +
                theme_bw() +
                theme(text = element_text(family = "serif"),
                        axis.text = element_text(size=15),
                        axis.title = element_text(size=18))

nov_plot <- df %>% ggplot(aes(y=novelty, x=date)) +
                geom_line(color = "steelblue") +
                labs(x = "Date", y = "Novelty") +
                theme_bw() +
                theme(text = element_text(family = "serif"),
                        axis.text = element_text(size=15),
                        axis.title = element_text(size=18))

patch <- res_plot / nov_plot
patch
# ggsave("fig/res_nov.png", patch, width=8)
```

# Resonance signal with changepoint periods and events
```{r}
plot<-ggplot()+
    # Resonance signal
    geom_line(data=df, 
              aes(date, resonance), 
              size=0.4)+

    # Background
    geom_rect(data = color_df,
              aes(xmin=as.Date(start),
                  xmax=as.Date(end),
                  ymin=-Inf,
                  ymax=Inf,
                  fill = as.factor(change_point)),
              alpha=0.1)+
    
    # Change point text
    geom_text(data=text_df,
              aes(x=x_pos, 
                  y=y_pos, 
                  label=str_wrap(text, width = 12), 
                  color=as.factor(change_point)), 
              size = 4,
              family="serif")+

    # Event points
    geom_point(data=covid_manual,
               aes(x=date,
                   y=resonance), 
               alpha=0.6,
               size = 3,
               color="sienna4")+
    
    # Text for event points
    geom_text(data=covid_manual,
              aes(x=date, 
                  y=y_pos,
                #   color=change_point,
                  label=n_event),
              size=3.5,
              family="serif",
              color="sienna4")+
    
    # Color and axis setup
    scale_fill_manual(values=fill_colors)+
    scale_color_manual(values=col_colors)+
    scale_x_date(breaks = floor_date(seq(min(df$date), max(df$date), by = 'quarter'), unit = "quarter"),
                 labels = date_format('%b-%Y'))+
    
    # Text size, font, axis labels and remove legend
    theme_bw()+
    labs(x="Date", y = "Resonance", title = "Resonance, change points and selected events")+
    theme(axis.text = element_text(size=11),
          axis.title = element_text(size=15),
          axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
          text = element_text(family = "serif"),
          legend.position = "none")

plot
# ggsave("fig/resonance_cp_events_2022.png", plot, width=8, height=5)
```

# Change point densities
```{r}
qs = c(0.03, 0.25, 0.5, 0.75, 0.97)

df2 <- df %>% group_by(change_point) %>%
    summarize(
              res_q = quantile(resonance, qs), 
              q = qs
              )

df2_lower <- df2 %>% filter(q == min(qs))
df2_upper <- df2 %>% filter(q == max(qs))

df2 <- data.frame(
    change_point = df2_lower$change_point,
    res_lower = df2_lower$res_q,
    res_upper = df2_upper$res_q
)

p <- ggplot(data = df) + 
    geom_density(aes(x = resonance), color='steelblue', size=1) +
    geom_text(data=df2_lower, mapping=aes(x=res_q, label=paste(round(res_q, 3)), y=60),  
                                #  angle=90, 
                                hjust = 'outward',
                                family = "serif",
                                size = 5
                                ) +
    geom_text(data=df2_upper, mapping=aes(x=res_q, label=paste(round(res_q, 3)), y=60),  
                                #  angle=90, 
                                hjust = 'inward',
                                family = "serif",
                                size = 5
                                ) +
    geom_segment(data = df2, mapping = aes(x=res_lower,xend=res_upper,y=0,yend=0), size=1) +
    facet_wrap(.~change_point) + 
    labs(y='Density', x = 'Resonance') +
    theme_bw() + 
    theme(strip.background =element_rect(fill="white"),
          axis.text = element_text(size=12),
          axis.title = element_text(size=15),
          strip.text.x = element_text(size = 15),
          text = element_text(family = "serif"))
p
# ggsave("fig/res_density.png", p, width=8, height=8)
```

# Resonance-novelty slope 
```{r}
for (n in unique(df$change_point)){
    df_changepoint <- df %>% filter(change_point==n)
    tmp_p <- ggplot(data = df_changepoint, aes(x = novelty, y = resonance)) + 
        geom_point(alpha=0.5) +
        geom_smooth(method = lm, color='steelblue') +  
        labs(title = paste("Change point period", n, sep = " "),
             x = "Novelty",
             y = "Resonance") +
        theme_bw() + 
        theme(text = element_text(family = "serif"),
              axis.text = element_text(size=11),
              axis.title = element_text(size=12),
              plot.title = element_text(size=15))
    eval(parse(text=paste('p_', n, ' <- tmp_p', sep='')))
}

patch <- (p_1 + p_2) / (p_3 + p_4) / (p_5 + p_6)
patch
# ggsave("fig/rn_slopes_2022_rm_outliers.png", patch, width=8, height=8)
```

# Correlation in change point windows
```{r}
labels = c('happiness',
            'trust',
            'expectation',
            'surprised',
            'anger',
            'contempt',
            'grief',
            'fear')
```

```{r}

for (cp in unique(df$change_point)){
    emos = df %>% subset(change_point == cp) %>% select(labels)
    corr_mat = melt(round(cor(emos),3))
    
    plot <- corr_mat %>% ggplot(aes(Var1, Var2, fill=value))+
                            geom_tile(color="White")+
                            scale_fill_gradient2(low="#234FAF",
                                                 high="#D75038", 
                                                 mid="#F4F0FF",
                                                 limit=c(-1,1),
                                                 midpoint=0,
                                                 name="")+
                            geom_text(aes(Var1, Var2, label = value), color = "black", size = 2) +
                            labs(title=paste("Change point period", cp))+
                            theme(axis.text.x = element_text(angle=20, hjust=1, size = 7),
                                  axis.text.y = element_text(size = 7),
                                  axis.title.x = element_blank(),
                                  axis.title.y = element_blank(),
                                  panel.grid.major = element_blank(),
                                  panel.border = element_blank(),
                                  panel.background = element_blank(),
                                #   axis.ticks = element_blank(),
                                  plot.title=element_text(size=8),
                                  text=element_text(family="serif", size=9))+
                            coord_fixed()
    eval(parse(text=paste('p_', cp, ' <- plot', sep='')))
    # ggsave(paste("fig/corr_mat_cp_", cp, ".png", sep = ""), plot, width=6, height=6)
}
patch = (p_1 + p_2) / (p_3 + p_4) / (p_5 + p_6)
patch
better_patch <- (patch & theme(legend.position = "right")) + plot_layout(guides = "collect")
better_patch

# ggsave("fig/corr_mat_2022.png", better_patch, width=7, height=7)
```


## Raw emotion distributions
```{r}
labels = c('happiness',
            'trust',
            'expectation',
            'surprised',
            'anger',
            'contempt',
            'grief',
            'fear')
```

Prepare df in long format
```{r}
emos <- df %>% select(c("date", labels))
emos_melt <- melt(emos, id=c("date"))

```


```{r}
raw_emotions <- emos_melt %>% ggplot(aes(date, value, color=variable))+
                geom_line()+
                theme_bw()+
                # scale_color_manual("Emotion", values = vals)+
                scale_color_brewer("Emotion", palette="Dark2")+
                labs(x="Date", y="Probability")+
                scale_x_date(
                labels = date_format('%b-%Y'))+
                theme(axis.text = element_text(size=11),
                    axis.title = element_text(size=15),
                    text = element_text(family = "serif", size=13)
                    )
raw_emotions
# ggsave("fig/raw_emotions_2022.png", raw_emotions, width=8, height=5)

```

